Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2018-01-25 paper/aldebaran.tex
phrasing
reference to sato

The stellar oscillations are best fit with a single, low-quality ($Q =
\CombinedQRange{}$) oscillatory mode with $\numax = 2.238^{+ 0.086}_{- 0.080}$
\muHz.  We find no improvement in the marginal likelihood (evidence) or other
model-selection information criteria \citep{Gelman2013} from models with
additional oscillatory modes, so we conclude that the large spacing $\Delta\nu$
is not observable in this data set.  We have determined the stellar properties
of Aldebaran with the additional constraint of \numax determined from this
analysis.  The details of the stellar modelling are presented in Appendix
\ref{sec:mod}.  We have considered the impact of our additional \numax
constraint and run our analysis for different sets of spectroscopic estimates.
For final values we adopt the \citet{2012Sheffield} spectroscopic solution as
being `middle of the road' estimates.  For the analysis without constraint on
\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
\, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the
inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  It is clear that the addition of \numax
provides substantially more precise estimates of mass and age. With this stellar
mass, the radial velocity signal translates to an inclination-dependent
planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$
\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
making it a possible low-mass brown dwarf candidate. A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran. We remain convinced that Aldebaran~b is a bona fide planet for reasons outlined in Appendix~\ref{gamma_dra}.

@@

From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters. Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis. Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements. One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}. The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars. Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case. This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible.

>>>>

The stellar oscillations are best fit with a single, low-quality ($Q =
\CombinedQRange{}$) oscillatory mode with $\numax = 2.238^{+ 0.086}_{- 0.080}$
\muHz.  We find no improvement in the marginal likelihood (evidence) or other
model-selection information criteria \citep{Gelman2013} from models with
additional oscillatory modes, so we conclude that the large spacing $\Delta\nu$
is not observable in this data set.  We have determined the stellar properties
of Aldebaran with the additional constraint of \numax determined from this
analysis.  The details of the stellar modelling are presented in Appendix
\ref{sec:mod}.  We have considered the impact of our additional \numax
constraint and run our analysis for different sets of spectroscopic estimates.
For final values we adopt the \citet{2012Sheffield} spectroscopic solution as
being `middle of the road' estimates.  For the analysis without constraint on
\numax we find estimates of the stellar properties as $M = 1.27^{+0.24}_{-0.2}
\, \mathrm{M_{\odot}}$ and age $4.86^{+3.56}_{-2.04} \, \rm Gyr$.  With the
inclusion of \numax we find $M = 1.16^{+0.07}_{-0.07} \, \mathrm{M_{\odot}}$ and
age $6.38^{+1.42}_{-1.12} \, \rm Gyr$.  It is clear that the addition of \numax
provides substantially more precise estimates of mass and age. With this stellar
mass, the radial velocity signal translates to an inclination-dependent
planetary mass of $M\,\sin\,{i} = 5.81 \pm 0.72~M_J$
\citep[Equation~1]{Torres2008}, a massive giant planet; for sufficiently high
inclinations ($\gtrsim 45^\circ$), the mass of Aldebaran~b may exceed $13~M_J$,
making it a possible low-mass brown dwarf candidate. A recent study by \citet{2018arXiv180105239H} has cast doubt on the validity of the detection of the massive planet $\gamma$~Draconis~b, and by extension similar planets around similar K~giants such as Aldebaran. We remain convinced that the $\gamma$~Draconis$ problem is not an issue in this case, and Aldebaran~b is a bona fide planet, for reasons outlined in Appendix~\ref{gamma_dra}.

@@

From this pilot study with Aldebaran, have shown that with limited quantities of data of limited quality, we can nevertheless do asteroseismology from the ground and obtain very precise estimates of stellar parameters. Furthermore, improvements in the algorithm may also permit the detection of not only~\numax but also~\Dnu from similar measurements, permitting more sophisticated asteroseismic analysis. Sufficient radial velocity data either already exist, or are trivial to obtain, in order to do this for essentially all bright giants; while for solar-like stars, these new methods will allow significantly relaxed observing requirements. One rich archive is the series of observations with the Hamilton \'{E}chelle Spectrograph at Lick Observatory, studying 373 G~and K~giants \citep[e.g.][]{frink2001,hekker2006,ortiz2016}. The method may also be useful in detecting rotational or other periodic variability in the LCES HIRES/Keck Precision Radial Velocity Exoplanet Survey \citep{butler17} of 1,624 FGKM dwarfs, although the time sampling is by design too coarse to detect $p$-mode oscillations in these stars. Furthermore, CARMA models and related methods will enhance deep, all-sky, sparse photometric surveys: an immediate future test of this will be from \hipparcos; while only~58 epochs of photometry are available for Aldebaran, at a sampling we find to be insufficient for our purposes, stars at higher latitudes may often have 150--200 epochs and more even sampling \citep{hipparcos_phot}, and the~14 bright~K giants in \hipparcos noted by \citet{bedding2000} are an ideal first test case. This can naturally be extended to \emph{Gaia} in space \citep{gaia}, or \emph{LSST} from the ground \citep{dmt,lsst,lsstbook}, from which many thousands of new asteroseismic determinations will be possible. In future RV surveys, it may even be possible to beat the \citet{2005PASJ...57...97S} $\sim 30$~m/s RV precision limit for red giants by taking many closely spaced observations and modelling-out the effects of stellar oscillations: this will allow us to dig deeper into this intrinsic stellar noise to detect less-massive planets around these stars.
